% Chapter 7

\chapter{Symmetries, topology and edge states } % Main chapter title

\label{Chapter7} % For referencing the chapter elsewhere, use \ref{Chapter7} 

\lhead{Part II. \emph{One wire}}
\chead{Chapter 7. \emph{Symmetries, topology \& edge states}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
In this chapter we study the topological effects. The main result is, that we show the ground state to be topologically non-trivial. In section \ref{sec.SymmetriesTRandPH} we define the time reversal, particle-hole and sublattice transformations. These we will use to come with a topological classification of the now developed Kitaev model system. This is done in section \ref{sec.Topology10foldway}. We then compute the topological index in two ways, respectively in section \ref{sec.topindexandphases} and \ref{sec.CS1}. These describe whether the system is in a topological or trivial phase. Finally, we will see what physical consequences the phases have by discussing the bulk-edge correspondence principle and edge states. An approximate solution for the edge states is found in section \ref{sec.edgestates}. The treatment is heavily based on the articles \cite{Ludwig.Topology, Chiu.Topology, Alicea}. 

\section{Time reversal, particle-hole and sublattice transformations}
\label{sec.SymmetriesTRandPH}
In this section we define the time reversal and particle-hole transformations. Let us remind ourselves of the form of the noninteracting Hamiltonian obtained after the mean field approximation of chapter \ref{Chapter4}: 
\begin{equation}
H_{FF} = -\sum_k \Delta_k \braket {c^\dagger_k c^\dagger_{-k}} + \frac{1}{2}\sum_{k} \begin{bmatrix} c^\dagger_k & c_{-k} \end{bmatrix} \begin{bmatrix} \varepsilon_k & \Delta_k \\ \Delta^*_k & -\varepsilon_k \end{bmatrix} \begin{bmatrix} c_k \\ c^\dagger_{-k} \end{bmatrix}. \nonumber 
\end{equation}

It is clear, that the Hamiltonian is translationally invariant. Explicitly, the (unitary) translation operator is defined through: $\tau(x_0)\psi_F(x)\tau^\dagger (x_0) = \psi_F(x+x_0)$. In momentum space, we hereby get $\tau(x_0) c_k \tau^\dagger(x_0) = \text{e}^{ikx_0}c_k$, as one might expect. Hence, terms like $c^\dagger_k c_k, c_k c^\dagger_k, c^\dagger_kc^\dagger_{-k}$ and $c_{-k}c_k$ are invariant under the translation. This means, that $H_{FF}$ is invariant as well. Because of the translational invariance, we \textit{inforce} that the time reversal and particle-hole transformations do not mix states with different positions. Specifically:
\begin{align}
T \begin{bmatrix} \psi_F(x) \\ \psi^\dagger_F(x) \end{bmatrix} T^{-1} &= U^\dagger_T \begin{bmatrix} \psi_F(x) \\ \psi^\dagger_F(x) \end{bmatrix}, \hspace{0.5cm} TiT^{-1} = -i, \nonumber \\
C \begin{bmatrix} \psi_F(x) \\ \psi^\dagger_F(x) \end{bmatrix} C^{-1} &= (U^*_C)^\dagger \begin{bmatrix} \psi^\dagger_F(x) \\ \psi_F(x) \end{bmatrix}, \hspace{0.5cm} CiC^{-1} = i.  
\label{eq.TRandPH.realspace}
\end{align}
Here $T$ and $C$ are the time reversal and particle-hole transformation respectively. $U_T$ and $U_C$ are $2\times 2$ matrices. Inforcing that transformed operators are fermionic as well, means that the matrices have to be unitary. The operation on $i$ specifies, that $T$ is antiunitary and $C$ is unitary. In momentum space this definition leads to:
\begin{align}
T \begin{bmatrix} c_k \\ c^\dagger_{-k} \end{bmatrix} T^{-1} &= U^\dagger_T \begin{bmatrix} c_{-k} \\ c^\dagger_{k} \end{bmatrix}, \nonumber \\
C \begin{bmatrix} c_k \\ c^\dagger_{-k} \end{bmatrix} C^{-1} &= (U^*_C)^\dagger \begin{bmatrix} c^\dagger_{-k} \\ c_{k} \end{bmatrix}. 
\label{eq.TRandPH.momentumspace}
\end{align}
By inspecting the transformations $TH_{FF}T^{-1}$ and $CH_{FF}C^{-1}$, we get the following symmetry requirements:
\begin{align}
TH_{FF}T^{-1} = H_{FF} \Leftrightarrow U_T\mathcal{H}^*_{FF,-k} U^\dagger_T = + \mathcal{H}_{FF,+k}, \nonumber \\
CH_{FF}T^{-1} = H_{FF} \Leftrightarrow U_C\mathcal{H}^*_{FF,-k} U^\dagger_C = - \mathcal{H}_{FF,+k}. 
\label{eq.Symmetryrequirements}
\end{align}
This means, that we can think of the second quantization transformations $T$ and $C$ in terms of first quantization antiunitary transformations $\mathcal{T} = U_TK$ and $\mathcal{C} = U_CK$, where $K$ is the complex conjugation operator. This is a general property of these transformation, not restricted to our specific system. See e.g. the articles \cite{Ludwig.Topology, Chiu.Topology}. The Hamiltonian at hand is a socalled Bogoliubov-de Gennes (BdG) Hamiltonian. These BdG Hamiltonians in general has a particle-hole symmetry. The reason is, that there is a redundancy in the matrix structure of the Hamiltonian. Explicitly, the structure contains a $2\times 2$ matrix, even though there is only one energy solution. Another way of putting this, is that the two Nambu spinors on each side of the kernel $\mathcal{H}_{FF,k}$ are not independent. We can transform one into the other by going to $-k$ and flipping the entries. The redundancy is here especially evident, since we can simply choose $C$ to have no effect on the Nambu spinor:
\begin{equation}
C \begin{bmatrix} c_k \\ c^\dagger_{-k} \end{bmatrix} C^{-1} =  \tau_1 \begin{bmatrix} c^\dagger_{-k} \\ c_{k} \end{bmatrix} = \begin{bmatrix} c_k \\ c^\dagger_{-k} \end{bmatrix}, 
\end{equation}
with $\tau_1$ the first Pauli matrix in particle-hole space. This of course means, that $H_{FF}$ is invariant under $C$. Since, it stems from a redundancy in the \textit{structure} of the Hamiltonian, it is often referred to as a particle-hole \textit{constraint} of BdG systems rather than a symmetry. This also means, that for the system to be consistent, we need $\tau_1\mathcal{H}^*_{FF,+k} \tau_1 = - \mathcal{H}_{FF,-k}$, which can also explicitly be checked. 

The time reversal case is somewhat trickier. Firstly, in general the pairing $\Delta_k$ is complex. However, as seen in the previous chapters we can explicitly find a real solution. This is also evident mathematically. For this purpose we let $\Delta_k \to \text{e}^{i\phi}\Delta_k$, with $\phi$ the phase of the pairing, and now $\Delta_k$ real. We then perform a gauge transformation in the $c_k$ operators according to: $c_k \to \text{e}^{-i\phi/2} c_k$. Since $\Delta_k$ depends linearly on $\braket{c_kc_{-k}}$ the resulting transformation of the pairing is:
\begin{equation}
\text{e}^{i\phi}\Delta_k \to \text{e}^{i\phi}\Delta_k\text{e}^{-i\phi} = \Delta_k. \nonumber
\end{equation}
With this the kernel $\mathcal{H}_{FF,k}$ is real, and with $U_T = i\tau_3$ we get a time reversal symmetry in accordance with equation \ref{eq.Symmetryrequirements}. Explicitly:
\begin{equation}
U_T\mathcal{H}^*_{FF,-k}U^\dagger_T = \tau_3\mathcal{H}^*_{FF,-k}\tau_3 = \begin{bmatrix} 1 & 0 \\ 0 & -1 \end{bmatrix}\begin{bmatrix} \varepsilon_k & -\Delta_k \\ -\Delta_k & -\varepsilon_k \end{bmatrix} \begin{bmatrix} 1 & 0 \\ 0 & -1 \end{bmatrix} = \begin{bmatrix} \varepsilon_k & \Delta_k \\ \Delta_k & -\varepsilon_k \end{bmatrix} = \mathcal{H}_{FF,k}, \nonumber
\end{equation}
since $\Delta_k$ is odd in $k$ and now also real. It is evident, that the found time reversal symmetry physically transforms a particle into a particle with opposite momentum:
\begin{equation}
T \begin{bmatrix} c_k \\ c^\dagger_{-k} \end{bmatrix} T^{-1} = \begin{bmatrix} i c_{-k} \\ - i c^\dagger_{k} \end{bmatrix}. \nonumber
\end{equation}
It is intuitively reasonable, that this \textit{should} be a symmetry of the system. Physically the fermions are bound in Cooper pairs of momentum $k$ and $-k$. This time reversal transformation transforms such a pair by flipping the momentum of both fermions. Explicitly: $Tc^\dagger_{-k}c^\dagger_{k}T^{-1} = -c^\dagger_{k}c^\dagger_{-k} = c^\dagger_{-k}c^\dagger_{k}$. Hereby the Cooper pair is invariant under the time reversal symmetry. Notice that this sounds a bit like inversion symmetry. Before performing the mean field approximation we can easily find an inversion symmetry of the system.\footnote{We can simply define $P\psi_F(x)P = \psi_F(-x)$.} However, in appendix \ref{AppendixB} we show that the mean field Hamiltonian has no inversion symmetry whatsoever. So the presence of the above time reversal symmetry is \textit{not} due to an underlying (unitary) inversion symmetry.

By composing the time reversal and particle-hole transformations we can form a third transformation: the sublattice (or chiral) symmetry $S = TC$. It is evident, that this transformation is antiunitary like $T$. The same analysis as in the above leads to the symmetry condition:
\begin{equation}
SH_{FF}S^{-1} = H_{FF} \Leftrightarrow U_S\mathcal{H}_{FF,+k} U^\dagger_S = - \mathcal{H}_{FF,+k}.
\end{equation}
Hence, the transformation in first quantization is unitary, but has to anticommute with the Hamiltonian. It is also evident, that since the system at hand both has a time reversal and particle-hole symmetry, we can simply form the product to get the symmetry with $U_S = i\tau_3\tau_1 = -\tau_2$. We explain the reason for including this third transformation in the next section. 

\section{Topology and the 10-fold way} \label{sec.Topology10foldway}
In the articles \cite{Ludwig.Topology, Chiu.Topology} it is studied, how one in the broadest possible sense can classify \textit{all} noninteracting fermionic Hamiltonians, that has an energy gap in the spectrum. This includes insulators and superconductors. Hence, the authors build a framework in which two distinct Hamiltonians are viewed as equivalent, if one can continuously deform one to the other without closing the energy gap. These sort of deformations are linked to the study of topological spaces, also called manifolds. Hence, the term \textit{topology}. Later we will see, that the topology of the Hamiltonian has geometrical effects in real space, the edge states. However, as the above makes it clear, this is not the source of the term topology.  

It turns out that one can use the three transformations of the previous section to classify the system within the above framework. Further, there are three possibilities of the time reversal and particle-hole symmetry. Either there is no symmetry, indicated by a $0$, or there is a symmetry and then the transformations can square to plus and minus the identity, indicated with $\pm 1$. The sublattice symmetry can either by absent, $0$, or present, $1$. Further, as the system at hand explicitly shows, if there is both a time reversal, $T$, and a particle-hole symmetry, $C$, present, there is also a sublattice symmetry, $S=TC$. Further, if only $T$ or $C$ is present of the two, there cannot be a sublattice symmetry. Else, we would be able to make the remaining symmetry by composition. However, in the case of no $T$ nor $C$ symmetry, there is still two possibilities for $S$: symmetry or no symmetry. This is the reason why this last symmetry is included. In total we then have $(3\cdot 3 - 1) + 2 = 10$ possibilities of combining the symmetries. It is therefore also referred to as the 10-fold way. 

Classifying the Hamiltonians in this framework puts the Hamiltonian in a specific Cartan class. The information from the classification is, that the socalled topological index has to be found in a specific set of numbers. This set can for example be the integers $\mathbb{Z}$. The result of the articles is then, that if two ground states have different integers as their topological index, they cannot be deformed continously into each other without closing the energy gap. The socalled periodic table for topological insulators and superconductors is shown in table \ref{tab.PeriodicTableTISC}.

\begin{table}[htb]
\centering
\caption{\textit{Periodic table of topological superconductors and insulators. $d$ is the spatial dimension. A 0-value for T,C and S indicates no symmetry. For T and C the sign indicates the sign of the square $T^2$ and $C^2$ for a symmetry present. $S=1$ indicates presence of a sublattice symmetry. The sets to the right indicates in what set of numbers the topological index should be found. }}
\begin{tabular}{|l|l l l|l l l l|}
\hline Cartan/$d$   &  T &  C & S					& 0 & 1 & 2 & 3 \\
\hline A    		&  0 &  0 & 0					& $\mathbb{Z}$ & 0 & $\mathbb{Z}$ & 0   			 \\
\hline AIII 		&  0 &  0 & 1					& 0 & $\mathbb{Z}$ & 0 & $\mathbb{Z}$   			 \\
\hline AI   		& +1 &  0 & 0					& $\mathbb{Z}$ & 0 & 0 & 0 			    			 \\
\hline BDI	       	& +1 & +1 & 1 					& $\mathbb{Z}_2$ & $\mathbb{Z}$ & 0 & 0 			 \\
\hline D	       	&  0 & +1 & 0 					& $\mathbb{Z}_2$ & $\mathbb{Z}_2$ & $\mathbb{Z}$ & 0 \\
\hline DIII	       	& -1 & +1 & 1 					& 0 & $\mathbb{Z}_2$ & $\mathbb{Z}_2$ & $\mathbb{Z}$ \\
\hline AII	       	& -1 &  0 & 0 				 	& $2\mathbb{Z}$ & 0 & $\mathbb{Z}_2$ & $\mathbb{Z}$  \\
\hline CII	       	& -1 & -1 & 1 					& 0 & $2\mathbb{Z}$ & 0 & $\mathbb{Z}_2$  			 \\
\hline C	       	&  0 & -1 & 0 					& 0 & 0 & $2\mathbb{Z}$ & 0  						 \\
\hline CI	       	& +1 & -1 & 1 					& 0 & 0 & 0 & $2\mathbb{Z}$  						 \\
\hline 
\end{tabular}
\label{tab.PeriodicTableTISC}
\end{table}

For the system at hand we have $\mathcal{T}^2 = \mathcal{C}^2 = \mathbb{I}$. This puts the system in the Cartan class BDI as table \ref{tab.PeriodicTableTISC} shows. This also means, that we in principle should look in the integers, $\mathbb{Z}$, to find the topological index of the system. 

\section{Topological index and phases}
\label{sec.topindexandphases}
In this section we will anticipate a potential breaking of time reversal symmetry. This means, that we only want to look at topological effects protected by the built-in particle-hole symmetry. In this sense, we compute the topological index corresponding to symmetry class $D$. We therefore search for a topological index in $\mathbb{Z}_2=\{-1,1\}$. This mimics the approach in \cite{Alicea}. The idea is the following. Define the vector $\mathbf{h}(k)$ with the components $h_z(k) = \varepsilon_k, h_y(k) = 0, h_x(k) = \Delta_k$. Since $\Delta_k$ has been chosen real, we can then see, that $\mathcal{H}_{FF,k} = \mathbf{h}(k)\cdot\boldsymbol\tau$, with $\boldsymbol\tau = (\tau_1,\tau_2,\tau_3)$ containing the Pauli matrices. The energy dispersion is $E_{F,k} = \sqrt{\varepsilon^2_k + \Delta^2_k} = |\mathbf{h}(k)|$. We hereby define a unit vector:
\begin{equation}
\hat{h}(k) = \frac{\mathbf{h}(k)}{|\mathbf{h}(k)|}. 
\label{eq.hhatdefinition}
\end{equation}


\begin{figure}
\center
\begin{tikzpicture}
\draw[-latex, thick] ( 0.0, -4) -- (0, 4) node[above]{$z$};
\draw[-latex, thick] (-0.1,  0) -- (4, 0) node[right]{$x$};
\node at (-0.4, 0) {$0$};

\draw[red, ultra thick] (0, -3.5) arc (-90:90:3.5);
\draw[-latex, red, ultra thick] (2.4648737348, -2.4848737348) -- (2.4748737348, - 2.4748737348);

\draw[domain=33.540844097:90,smooth,variable=\x,blue, thick] plot ({3.5 * cos(\x)},{3.5 * sin(\x)});
\draw[-latex, blue, thick] (0.6853426578, 3.4322601227) -- (0.6953426578, 3.4302330223);
\draw[-latex, blue, thick] (2.4748737348,  2.4748737348) -- (2.4648737348, 2.4848737348);

\end{tikzpicture}
\caption{In red (blue) $(h_x(k),h_z(k))$ is plotted as function of $k\in \{0, \infty\}$ for $\mu > 0$ ($\mu < 0$). Notice that for $\mu < 0$ the curve starts and ends at the north pole.}
\label{fig.hhatplot}
\end{figure}

It is then clear, that as long as there is an energy gap, $\hat{h}(k)$ is well defined. Since $E_{F,k=0} = |\mu|$, we see that the energy is gapped exactly when $\mu \neq 0$. Finally, since $\Delta_{k=0} = 0$ and $\Delta_k \overset{k\to \infty}{\to} 0$, we notice that: 
\begin{equation}
\hat{h}(k=0) = -\text{sgn}(\mu)\hat{z} = p_0\hat{z}, \hspace{0.5cm} \hat{h}(k=\infty) = \hat{z} = p_{\infty}\hat{z},
\label{eq.hhatk0kinfty}
\end{equation}
where $\text{sgn}$ denotes the sign function. The idea is then, that we can specify a number $\nu \in \{-1,1\}$ by:
\begin{equation}
\nu = p_0p_{\infty} = -\text{sgn}(\mu).
\label{eq.topinvnudefition}
\end{equation}
The claim is, that we have hereby defined a topological index, which specifies two states, that cannot be deformed into each other without closing the energy gap. Graphically this is clear from figure \ref{fig.hhatplot}. In red (blue) I have plotted the behaviour of $\hat{h}$ for $\mu > 0$ ($\mu < 0$), omitting the zero component $h_y$. For $\mu > 0$ we see, that we go from the south pole to the north pole as $k$ goes from $0$ to $\infty$. For $\mu < 0$ on the other hand, we see that $\hat{h}(k)$ starts out at the north pole, then goes somewhat along the circle arc before returning to the north pole. This is of course the same behaviour as written in equation \eqref{eq.hhatk0kinfty}. It is then clear, that there is no way of continuously deforming one map to the other. Further, we notice that the mapping becomes ill-defined exactly when $\mu = 0$. This is also evident from the appearance of $\text{sgn}(\mu)$ in $\nu$. We notice, that since $\mu(T) \to -\infty$ and $\Delta_k(T)\to 0$ for $T\to \infty$, the $\mu < 0$ phase is connected continously to the normal phase. Hence, the normal phase, and thereby also the vacuum, is associated with $\nu = 1$. 

For $T = 0$ the normal phase has $\mu = \epsilon_{F,0} > 0$. We can hereby distinguish between two regimes. The first is the weak coupling regime. The pairing does not significantly alter the chemical potential: $\mu(T=0) > 0$. Hereby $\nu = -1$ and the ground state is topologically \textit{non}-trivial. The second is the strong coupling regime, where $\mu(T = 0) < 0$, $\nu = 1$ and the ground state is topologically trivial. This is in complete harmony with the findings in \cite{Alicea} for the Kitaev chain.\footnote{In the article by Alicea the fermions sit in a lattice, but the physical properties are essentially the same.}

We are only investigating the weak coupling regime, even with $\mu \approx \epsilon_{F,0}$. From the above arguments, this means, that we have found a ground state with $\nu = -1$. So far, we have investigated the bulk properties of the superfluid by imposing cyclic boundary conditions. Imagine it now as a wire with open ends. This means, that there are junctions at the ends between a $\nu = -1$ (the superfluid wire) and a $\nu = 1$ state (the vacuum). It also means, that when we spatially follow the transition between the two phases, the wire and the vacuum, $\nu$ has to change its value abrubtly at the boundary. Hence, the energy gap has to dramatically drop to $0$ at the boundary. 

The interesting question is now: are there states that are topologically protected? More precisely, we are asking, whether there exist low energy states, that can only be removed by breaking the symmetries of the system. The answer lies in the built-in particle-hole symmetry of the Hamiltonian. As we saw in section \ref{sec.SymmetriesTRandPH} we have a particle-hole symmetry $\mathcal{C} = \tau_1 K$, with $K$ the complex conjugation operator. Hereby $\mathcal{C}\mathcal{H}_{FF,k} = -\mathcal{H}_{FF,-k}\mathcal{C}$. Let us assume, that we have a single particle energy solution in $k$: $\mathcal{H}_{FF,k}\ket{\psi_k} = E_k\ket{\psi_k}$. We hereby directly get:
\begin{equation}
E_k\mathcal{C}\ket{\psi_k} = \mathcal{C}\mathcal{H}_{FF,k}\ket{\psi_k} = -\mathcal{H}_{FF,-k}\mathcal{C}\ket{\psi_k} \Rightarrow \mathcal{H}_{FF,-k}\left(\mathcal{C}\ket{\psi_k}\right) = -E_k\left(\mathcal{C}\ket{\psi_k}\right).
\end{equation}
This shows, that if $\ket{\psi_k}$ has the energy $E_k$, then $\mathcal{C}\ket{\psi_k}$ has the energy $-E_k$. In other words for every positive energy solution there is also a negative one, as is shown to the left in figure \ref{fig.edgestates}. The bulk spectrum is here indicated by a solid line above the bulk gap $\min_k[E_{F,k}]$ and mirrored in $E = 0$. Now suppose that we introduce a perturbation, that respects the particle-hole symmetry. The energies therefore still come in plus/minus pairs, but apart from that the energies can change. Hence, we can perturb any low energy state with $E \neq 0 $ away. This is indicated in the figure by the black dots and arrows. However, the \textit{single} state at $E = 0$ remains exactly there, because if it moved either up or down, there would not be the plus/minus symmetry in the energies. This leads to an energy spectrum, which is \textit{robust} against symmetry-conserving perturbations with the continuum above the bulk gap and a \textit{single} zero energy edge state as indicated in figure \ref{fig.edgestates}. 

With no energy cost edge states at the ends of the wire can hereby emerge and are robust to symmetry-conserving perturbations. This phenomenon is known as the bulk-edge correspondence principle, since it connects bulk effects, $\nu = -1$, to an edge effect (the edge states). As is indicated to the right in figure \ref{fig.edgestates}, the state has to be localised on the boundaries. The reason is, that all bulk states are already occupied, and so a zero energy bulk state is not available. 

\begin{figure}
\center
\begin{tikzpicture}
\draw[|-latex, thick] (0, 0) -- (0,  2) node[above]{$E$};
\draw[-, thick]   (0, 0) -- (0, -2);
\node at (-0.4, 0) {$0$};

\draw[-, dashed] (0, 1)--(1, 1);
\draw[-, thick] (-0.1, 1)--(0.1, 1);
\node at (-1.17,1) {$\min_k[E_{F,k}]$};

\draw[-, dashed] (0, -1)--(1, -1);
\draw[-, thick] (-0.1, -1)--(0.1, -1);
\node at (-1.35,-1) {$-\min_k[E_{F,k}]$};

\draw[-, ultra thick] (1, 1)--(1, 2);
\draw[-, ultra thick] (1, -1)--(1, -2);
\node[red] at (1, 0) {\textbullet};

\node at (1, 0.5) {\textbullet};
\draw[-latex] (1, 0.5) --  (1,  1);
\node at (1, -0.5) {\textbullet};
\draw[-latex] (1, -0.5) -- (1, -1);

\draw[-, dashed] (0,0.01) --(1,0.01);

\draw[scale=0.5,domain=-1:1,smooth,variable=\x,red, thick] plot ({\x + 6},{ 3 * exp{- 10 * \x*\x}});
\draw[scale=0.5,domain=-1:1,smooth,variable=\x,red, thick] plot ({\x + 16},{ 3 * exp{- 10 * \x*\x}});

\node at (3, -0.4) {$0$};
\node at (8, -0.4) {$\mathcal{L}$};
\draw[|-latex, thick ] (3,0) -- (9,0) node[right]{$x$};

\draw[-, ultra thick ] (3,0) -- (8,0);
\end{tikzpicture}
\caption{To the left the energy spectrum is sketched. Because of the particle-hole symmetry, there is a negative energy state for every positive. There is a continuum of states above $\min_k[E_{F,k}]$ and below $-\min_k[E_{F,k}]$. At the boundary of the wire the gap closes, but the particle-hole symmetry only protects the single state at $E = 0$, indicated by the red dot. The corresponding wave function is sketched to the right.}
\label{fig.edgestates}
\end{figure}

\section{Chern-Simons invariant, Wilson loop and winding number}
\label{sec.CS1}
In this section we calculate the Chern-Simons topological invariant of the system and in turn the gauge independent Wilson loop. It is based on the article \cite{Ryu.Topology}. We do not anticipate any breaking of time reversal symmetry. The topological index is therefore to be found in the integers, $\mathbb{Z}$. After this rather abstract calculation we will link the Chern-Simons invariant to the winding number of a specific mapping. We will work with $\Delta_k$ real. 

The Chern-Simons invariant is generally a half integer: $\dots, -3/2, -1, -1/2, 0, 1/2, \dots$. The topological index, generally called the winding number, is twice the Chern-Simons invariant, hence an integer. This gives us the desired topological index in $\mathbb{Z}$. We will return to this winding number at the end of the section. In one spatial dimension the Chern-Simons invariant is defined as:
\begin{equation}
\text{CS}_1 = \frac{i}{2\pi}\int_{\text{BZ}}dk \; \tr\left[\mathcal{A}_k\right], \nonumber
\end{equation}
where BZ is short for the first Brillouin zone, and $\mathcal{A}_k$ is the socalled Berry connection. In the literature the Chern-Simons invariant is also referred to as the (total) polarisation. This naming has a physical reasoning in electronic topological insulators, because it is actually a measure of the charge polarisation as a consequence of the edge states \cite{FuKane2006}. Since there are no charge carriers in our system, we will not use this naming. However, in a similar manner the norm of the winding number $|\nu| = 2|\text{CS}_1|$ counts the number of edge states \cite{Chiu.Topology}. Hence, the invariant has a very specific physical interpretation. 

Now let $\ket{e^{-}_k}$ be the eigenvector to $\mathcal{H}_{FF,k}$ with eigenvalue $-E_{F,k}$. For a single occupied state, the Berry connection is defined as:
\begin{equation}
\mathcal{A}_k = \bra{e^{-}_k}\partial_k\ket{e^{-}_k}. \nonumber
\end{equation}
The result of Chern-Simons theory is then, that the Chern-Simons invariant is a topological invariant. By this we mean, that it characterizes different topological phases, that can only change when the energy gap closes. It turns out that the calculation is easier in a different basis, than the one we have been working in so far. In appendix \ref{AppendixC} we show, that $\tr\left[\mathcal{A}_k\right]$ is independent of basis as long as the basis change is independent of $k$. This simply comes down to the basis invariance of the trace operation $\tr[\dots]$. This is of course very reasonable; a topological invariant should not depend on the chosen basis. We use the following matrix to change basis:
\begin{equation}
U = \text{e}^{i\tau_1 \frac{3\pi}{4}} = \frac{1}{\sqrt{2}}\left(-\mathbb{I} + i\tau_1\right). \nonumber 
\end{equation}
This transformation amounts to letting $(\tau_1, \tau_2, \tau_3) \overset{U}{\to} (\tau_1, \tau_3, -\tau_2)$. Since $\mathcal{H}_{FF,k} = \Delta_k\tau_1 + \varepsilon_k\tau_3$, in the new basis we get: $\tilde{\mathcal{H}}_{FF,k} = U\mathcal{H}_{FF,k}U^\dagger = \Delta_k\tau_1 - \varepsilon_k\tau_2$. The reason for changing basis is, that the eigenvectors are simpler and continuous as long as the energy gap is nonzero. Explicitly:
\begin{equation}
\ket{\tilde{e}^{-}_k} = \frac{1}{\sqrt{2}}\begin{bmatrix} -\frac{\Delta_k + i \varepsilon_k}{E_{F,k}} \\ 1 \end{bmatrix}. \nonumber 
\end{equation}
With this and a bit of calculating the Berry connection becomes:
\begin{equation}
\mathcal{A}_k = \bra{\tilde{e}^{-}_k}\partial_k\ket{\tilde{e}^{-}_k} = -\frac{i}{2E^2_{F,k}}(\varepsilon_k\partial_k\Delta_k - \Delta_k \partial_k \varepsilon_k). \nonumber 
\end{equation}
Since the Brillouin zone is simply the entire axis, $k\in (-\infty, \infty)$, we get:
\begin{equation}
\text{CS}_1 = \frac{i}{2\pi}\int_{\text{BZ}} dk\; \text{tr}\left[\mathcal{A}_k\right] = \frac{1}{4\pi}\int_{-\infty}^{\infty} dk \;  \frac{\varepsilon_k\partial_k\Delta_k - \Delta_k \partial_k \varepsilon_k}{\varepsilon^2_k + \Delta^2_k} = \frac{1}{2\pi}\int_0^\infty dk \; \frac{\varepsilon_k\partial_k\Delta_k - \Delta_k \partial_k \varepsilon_k}{\varepsilon^2_k + \Delta^2_k}, 
\label{eq.CSinv1}
\end{equation}
where the last equality stems from the fact, that the integrand is even, since $\Delta_k$ is odd. The integrand has a primitive given by:
\begin{equation}
G_k(c) = \arctan\left(\frac{\Delta_k}{\varepsilon_k}\right) + c \nonumber
\end{equation}
For $\mu < 0$, $\varepsilon_k > 0$ for all $k$ and $G_k(c = 0)$ is continuous. In this case the invariant simply becomes:
\begin{equation}
\text{CS}_1(\mu < 0) = \left. \frac{1}{2\pi}\arctan\left(\frac{\Delta_k}{\varepsilon_k}\right)\right|_{0}^{\infty} = 0,
\end{equation}
since $\Delta_k$ is zero for $k = 0$ and $k = \infty$ and since $\varepsilon_k$ is nonzero in both limits. Since $\Delta_k = 0$ in the normal phase, we have $\text{CS}_1 = 0$ here as well. Hence, the system is topologically trivial for $\mu < 0$. 

For $\mu > 0$, $G_k(c = 0)$ suffers a discontinuity in $k_0$ defined by $\varepsilon_{k_0} = \frac{k_0^2}{2m_F} - \mu = 0$. Explicitly:  
\begin{align}
\lim_{k \uparrow +k_0} G_k(c = 0)   &\overset{\varepsilon_k \uparrow 0}{=}   -\frac{\pi}{2}\text{sgn}(\Delta_{k_0}), \nonumber \\
\lim_{k \downarrow +k_0} G_k(c = 0) &\overset{\varepsilon_k \downarrow 0}{=} +\frac{\pi}{2}\text{sgn}(\Delta_{k_0}), \nonumber
\end{align}
This follows from the fact, that $\arctan(x) \overset{x \to \pm \infty}{\to} \pm \frac{\pi}{2}$. To mend this discontinuity we let $ c = -\pi\text{sgn}(\Delta_{k_0})$ for $k > k_0$. Notice, that $\Delta_{k_0} \neq 0$, since else the energy gap would close in $k_0$ ($E_{F,k} = \sqrt{\varepsilon^2_k + |\Delta_k|^2}$). Hence, the sign of $\Delta_{k_0}$, $\text{sgn}(\Delta_{k_0})$, is well defined. Therefore, the integrand in this situation has the continuous integrand:
\begin{equation}
G_k = \left\{ \begin{matrix} 
\arctan\left(\frac{\Delta_k}{\varepsilon_k}\right) - \pi\text{sgn}(\Delta_{k_0}), & k > k_0, \\
\arctan\left(\frac{\Delta_k}{\varepsilon_k}\right), & 0 < k < k_0,
  \end{matrix} \right.
\label{eq.Gkmugreater0}
\end{equation}
Using this the invariant for $\mu > 0$ becomes:
\begin{align}
\text{CS}_1(\mu > 0) = \left. \frac{1}{2\pi} G_k\right|_0^{\infty} = -\frac{\text{sgn}(\Delta_{k_0})}{2}.
\label{eq.CSinv2}
\end{align}
Since $\Delta_k$ is real the only remaining gauge freedom is the sign of the $\Delta_k$: $\Delta_k \to - \Delta_k$. Hence, the invariants $-1/2$ and $+1/2$ are equivalent. This explicitly shows, that the Chern-Simons invariant depends on the chosen gauge. To account for this, we compute the \textit{gauge independent} Wilson loop $W_1$. We get: 
\begin{equation}
W_1 = \text{e}^{2\pi i \text{CS}_1} = -\text{sgn}(\mu). 
\end{equation}
This means, that $W_1 = -1$ is the (only) topological non-trivial phase. We also see, that the condition is exactly the same as the one found in the previous section. The current calculation does however have additional information. As mentioned above the norm of the winding number counts the number of edge states. Since $|\nu| = 1$ in the only non-trivial case this shows, that the gas only supports up to one edge state. It is not possible for the system at hand to produce more edge states than this. The edge states analysis of the previous section is therefore unaltered in the present context. Notice, that this does not depend on the details of the model; the only used assumptions are: $\Delta_k$ odd, $\varepsilon_k$ only zero in $k = k_0$ and the nonzero energy gap. 

The question now becomes, what winding the winding number is a measure of. Here we return to the mapping investigated in the previous section: 
\begin{equation}
\hat{h}(k) = \frac{\mathbf{h}(k)}{|\mathbf{h}(k)|}, \hspace{0.5cm} \mathbf{h}(k) = \begin{bmatrix} \Delta_k & 0 & \varepsilon_k \end{bmatrix}, \hspace{0.5cm}  |\mathbf{h}(k)| = E_{F,k} \nonumber
\end{equation}
This mapping is as already discussed well-defined as long as the energy gap is nonzero. Now let $\theta_k$ be the angle from the $z$ axis to $\hat{h}(k)$. We then ask how many times the mapping $\hat{h}(k)$ winds around $\mathbf{h} = 0$. For $k = \pm \infty$ we have $\hat{h}(k) = (0, 0, 1)$. Hence, $\hat{h}$ winds an integer number of times around zero. This integer is given by the winding number of the mapping:
\begin{equation}
w = \frac{1}{2\pi}\oint d\theta_k, \nonumber
\end{equation}  
where we integrate from the start to the end of the Brillouin zone: $k \in (-\infty, \infty)$. The angle $\theta_k$ is given by: $\tan(\theta_k) = \frac{h_x(k)}{h_z(k)} = \frac{\Delta_k}{\varepsilon_k}$. In turn $\theta_k = \arctan\left(\frac{\Delta_k}{\varepsilon_k}\right) + c$ for some constant $c$. Notice, that this is exactly the primitive $G_k(c)$ to the integrand in $\text{CS}_1$. Hence, the winding number of $\hat{h}$ becomes:
\begin{equation}
w = \frac{1}{2\pi}\oint d\theta_k = \frac{1}{2\pi}\int_{-\infty}^{\infty} dk \frac{\varepsilon_k\partial_k\Delta_k - \Delta_k\partial_k\varepsilon_k}{\varepsilon^2_k + \Delta^2_k} = 2\text{CS}_1 = \nu. \nonumber
\end{equation}
The winding number of the map $\hat{h}(k)$ around zero is exactly the winding number connected to the Chern-Simons invariant. From this we also see, that winding numbers of differing sign are equivalent, since it simply corresponds to which way around $0$ $\hat{h}$ winds. This corresponds to the gauge transformation $\Delta_k \to - \Delta_k$. Further, two winding numbers $w_1$ and $w_2$ are topological distinct if $|w_1|\neq |w_2|$ in the following sense. The map must wind an integer number of times around zero. Therefore, we cannot continuously deform the mapping resulting in $w_1$ to the mapping resulting in $w_2$, hence they are topolologically distinct in a mathematical sense. Summarising, when the BdG Hamiltonian has a time reversal symmetry squaring to plus the identity, the topologically distinct physical phases of the system correspond to topologically different winding numbers of $\hat{h}$.  

\section{Edge states}
\label{sec.edgestates}
In this section we will come with an approximate analytical solution for the edge states between topologically distinct phases. This is primarily to get an idea of what these edge states are. The analysis is largely based on \cite[pp. 196-198]{BernevigTITSC}. 

\begin{figure}
\center
\begin{tikzpicture}
\begin{axis}[samples=150, xtick=\empty, ytick=\empty, axis line style={draw=none}, 
    xmin=-1, xmax=6,
    ymin = -1/2, ymax = 1/2,
    axis lines=center,
    domain=-1:6,
    ]

    \addplot [mark=none,draw=blue, thick] {1/2 * (tanh(5 * \x) - tanh(5 * (\x - 5) ) - 1)};
\end{axis}


\draw[scale=0.5,domain=-1:1,smooth,variable=\x,red, thick] plot ({\x + 0.975 * 2},{ 2 * 2.85 + 3 * exp{- 10 * \x*\x}});
\draw[scale=0.5,domain=-1:1,smooth,variable=\x,red, thick] plot ({\x + 5.875 * 2},{ 2 * 2.85 + 3 * exp{- 10 * \x*\x}});

\node at (0.975 + 0.12, 2.85 - 0.3) {$0$};
\node at (5.875-0.12, 2.85 - 0.3) {$\mathcal{L}$};
\draw[-latex, thick ] (0,2.85) -- (7,2.85) node[right]{$x$};
\draw[-, ultra thick ] (0.975,2.85) -- (5.875,2.85);
\draw[|-, thick ] (0.975,2.85) -- (1,2.85);
\draw[-, ultra thick ] (0.975,2.85) -- (5.875,2.85);
\draw[-|, thick ] (5.7,2.85) -- (5.875,2.85);

\coordinate (a) at (7.8, 5.682);
\coordinate (anode) at (7.5, 5.682);
\coordinate (b) at (8.3, 5.682);
\coordinate (c) at (7.8, 5.182);
\coordinate (d) at (8.3, 5.182);
\coordinate (cnode) at (6.7, 5.182);

\draw[-, thick, blue] (a) -- (b);
\node at (anode) {$\mu$}; 
\draw[-, thick, red] (c) -- (d);
\node at (cnode) {Edge states}; 
\end{tikzpicture}
\caption{Red: Edge states. Blue: chemical potential $\mu = \mu(x)$. Due to the spatial variation of the chemical potential the fermions are confined to the interval between $x = 0$ and $x = \mathcal{L}$. }
\label{fig.edgestatesmux}
\end{figure}

The single wire system has the Hamiltonian:
\begin{equation}
H_{FF} = E_0 + \frac{1}{2}\sum_{k} \begin{bmatrix} c^\dagger_k & c_{-k} \end{bmatrix} \begin{bmatrix} \varepsilon_k & \Delta_k \\ \Delta_k & -\varepsilon_k \end{bmatrix} \begin{bmatrix} c_k \\ c^\dagger_{-k} \end{bmatrix},
\end{equation}
with energy $E_{F,k} = \sqrt{ \varepsilon^2_k + \Delta^2_k}$. We would like to describe this Hamiltonian in real space, because we now ask whether certain states exist at the ends of the wire. To do this we define the operator $\Delta(p)$ as a function of the momentum operator $p = -i\partial_x$ in the following manner:
\begin{equation}
\Delta(p)\text{e}^{ikx} = \Delta_k\text{e}^{ikx}.
\label{eq.Deltapdef}
\end{equation}
Hence, we replace the variable dependency of $k$ in $\Delta_k$ with an operator dependency on $p$ in $\Delta(p)$. Further, we drop the ground state grand energy $E_0$. With this definition we get the Hamiltonian in real space:
\begin{equation}
H_{FF} = \frac{1}{2}\int dx \begin{bmatrix} \psi^\dagger_{F}(x) & \psi_{F}(x) \end{bmatrix} \mathcal{H}_{FF}(x) \begin{bmatrix} \psi_{F}(x) \\ \psi^\dagger_{F}(x) \end{bmatrix}, \hspace{0.5cm} \mathcal{H}_{FF}(x) = \begin{bmatrix} \frac{p^2}{2m_F} - \mu  & \Delta(p) \\ \Delta(p) & -\left( \frac{p^2}{2m_F} - \mu \right) \end{bmatrix},
\label{eq.MFHamiltonianrealspace}
\end{equation}
For the sake of clarity let us see how e.g. the term $\Delta_kc_{-k}c_{k}$ appears from this expression:
\begin{align}
\int dx \; \psi_{F}(x) \Delta(p) \psi_{F}(x) &= \frac{1}{\mathcal{L}}\sum_{k,q}c_k c_q\int dx \; \text{e}^{ikx}\Delta(p)\text{e}^{iqx} = \frac{1}{\mathcal{L}}\sum_{k,q}\Delta_qc_kc_q\int dx \; \text{e}^{i(k + q)x} \nonumber \\
&= \sum_{k}\Delta_kc_{-k}c_k, \nonumber 
\end{align}
since $\int dx\; \text{e}^{i(k + q)x} =\mathcal{L} \delta_{k,-q} $. Here we use the Fourier decomposition $\psi_{F}(x) = \frac{1}{\sqrt{L}}\sum_k\text{e}^{ikx}c_k$. Now we form junctions of topogically distinct phases by changing the sign of $\mu$ at $x = 0$ and $x = \mathcal{L}$. Hence, we let $\mu = \mu(x)$ as shown in figure \ref{fig.edgestatesmux}. Due to the spatial variation of the chemical potential the fermions are confined to the wire between $x = 0$ and $x = \mathcal{L}$. In first quantization we find solutions by solving $\mathcal{H}(x)\psi(x) = E\psi(x)$. The edge states are zero energy modes, so we solve for $ E = 0$. We assume, that $\mu(x)$ varies slowly over the interparticle length scale $1/k_F$. When this is fulfilled the solution $\psi(x)$ will also vary slowly and hence the curvature of $\psi(x)$ proportional to $p^2\psi(x)$ is negligible. This means, that to a good approximation we only keep the operator $p$ up to first order. Since $\Delta_k$ is linear in $k$ for $k/k_F \ll 1$, then $\Delta(p)$ is linear in $p$, when $p/k_F$ used on a state is small. Hence, we ignore $p^2/2m_F$ and let $\Delta(p) = \Delta \cdot \tilde{p}$, with $\Delta = k_F\left.\frac{\partial \Delta_k}{\partial k}\right|_{k=0}$ and $\tilde{p} = p/k_F$. We assume $\Delta > 0$. We take the ansatz:
\begin{equation}
\psi_1(x) = \exp\left(-\frac{k_F}{\Delta}\int_{0}^{x} dx'\; \mu(x') \right)\begin{bmatrix} u_1 \\ u_2 \end{bmatrix}, \hspace{0.5cm} \psi_2(x) = \exp\left(+\frac{k_F}{\Delta}\int_{\mathcal{L}}^{x} dx'\; \mu(x') \right)\begin{bmatrix} v_1 \\ v_2 \end{bmatrix}, \nonumber
\end{equation}  
respectively at the lower and upper edge: $x = 0$ and $x = \mathcal{L}$. Since $\Delta$ is in units of energy the exponents are unitless. Since $\Delta > 0$, we see that these states are exponentially localized at the edges. From this we get two eigenvalue equations:
\begin{align}
0 &= \mathcal{H}_{FF}(x)\psi_1(x) = \begin{bmatrix} - \mu  & i\mu \\ i\mu & + \mu \end{bmatrix}\psi_1(x) \Rightarrow 0 = \begin{bmatrix} - \mu  & i\mu \\ i\mu & + \mu \end{bmatrix}\begin{bmatrix} u_1 \\ u_2 \end{bmatrix}, \nonumber \\
0 &= \mathcal{H}_{FF}(x)\psi_2(x) = \begin{bmatrix} - \mu  & -i\mu \\ -i\mu & + \mu \end{bmatrix}\psi_2(x) \Rightarrow 0 = \begin{bmatrix} - \mu  & -i\mu \\ -i\mu & + \mu \end{bmatrix}\begin{bmatrix} v_1 \\ v_2 \end{bmatrix} \nonumber.
\end{align}
By solving these equations we get:
\begin{equation}
\psi_1(x) = g_1(x)\frac{\text{e}^{i\pi/4}}{\sqrt{2}}\begin{bmatrix} 1 \\ -i \end{bmatrix}, \hspace{0.5cm} \psi_2(x) = g_2(x)\frac{\text{e}^{-i\pi/4}}{\sqrt{2}}\begin{bmatrix} 1 \\ +i \end{bmatrix}, \nonumber
\end{equation}
with $g_1(x) = \frac{1}{N}\text{e}^{-\frac{k_F}{\Delta}\int_{0}^{x} dx' \mu(x')}$, $g_2(x) = \frac{1}{N}\text{e}^{+\frac{k_F}{\Delta}\int_{\mathcal{L}}^{x} dx' \mu(x')}$ and $N^2 = \int dx |g_j(x)|^2$ a normalization constant. Notice the sign differences between the two solutions. The above eigenstates define the diagonalisation matrix $U(x) = \begin{bmatrix} \psi_1(x) & \psi_2(x) \end{bmatrix}$. This has the unitarity property, that $\int dx \; U^\dagger(x) U(x) = \mathbb{I}$. We then let the operators $\gamma_{0,1}, \gamma_{0,2}$ be given by:
\begin{equation}
\begin{bmatrix} \psi_F(x) \\ \psi_F^\dagger(x) \end{bmatrix} = U(x) \begin{bmatrix} \gamma_{0,1} \\ \gamma_{0,2} \end{bmatrix} \Rightarrow \begin{bmatrix} \gamma_{0,1} \\ \gamma_{0,2} \end{bmatrix} = \int dx \; U^\dagger(x) \begin{bmatrix} \psi_F(x) \\ \psi_F^\dagger(x) \end{bmatrix}.
\label{eq.Majoranaedgemodedef} 
\end{equation}
Written out the operators are:
\begin{equation}
\gamma_{0,1} = \int dx \; g_1(x) \frac{\text{e}^{-i\pi/4}}{\sqrt{2}}(\psi_F(x) + i\psi^\dagger_F(x)), \hspace{0.5cm} \gamma_{0,2} = \int dx \; g_2(x) \frac{\text{e}^{+i\pi/4}}{\sqrt{2}}(\psi_F(x) - i\psi^\dagger_F(x)).
\label{eq.Majoranaedgemodegeneraldef} 
\end{equation}
In this manner $\psi_j(x)$ and $\gamma_{0,j}\ket{\text{S}}_0$ are equivalent. These operators are special in the sense, that they are their own antiparticle: $\gamma_{0,j}^\dagger = \gamma_{0,j}$.\footnote{This also illuminates why we chose the rather arbitrary phase factors $\text{e}^{i\pi/4}$ and $\text{e}^{-i\pi/4}$. Had we \textit{not} done this, there would be some phase difference between $\gamma^\dagger_{0,j}$ and $\gamma_{0,j}$.} Such a fermionic particle is called a Majorana mode in condensed matter physics. Besides being their own antiparticle, they have to obey the modified anticommutator relation:
\begin{equation}
\{\gamma_{0,i}, \gamma_{0,j} \} = \delta_{i,j}. \nonumber
\end{equation}
This is explicitly verified to be the case for the above operators.\footnote{Often one also meets the requirement, that the anticommutator equals $2\delta_{i,j}$. This is simply a matter of convention.} From this we can form a regular but highly nonlocal fermionic operator: $d = \frac{1}{2}(\gamma_{0,1} + i\gamma_{0,2})$. The key result of this section is thus, that $\ket{\text{S}}_0$ and $d^\dagger\ket{\text{S}}_0$ are degenerate ground states of the system. The topologically non-trivial state is therefore characterized by the presence of \textit{one} more particle than the trivial one, which is exponentially located on the edges.\footnote{In the literature one also meets the notion of a Majorana mode at each end, hence two in total. However, since $\gamma^\dagger_0\gamma_0 = \gamma_0^2 = \frac{1}{2}$ for Majorana operators, we cannot count them. Hence, the notion of a specific number of Majorana modes is undefined.}

The question now becomes the following. Is it possible to make a simple physical change of the system, that allows us to observe a topological phase transition? The proposal is to put in a \textit{second} wire parallel to the one already studied. By doing this we get two kinds of interactions. The first is the \textit{intra}wire interaction, the interaction internally in each wire, which we have already studied in detail. The second is the \textit{inter}wire interaction, the interaction between the two wires. Further, we get an extra parameter, namely the distance $d$ between the wires. It will allow us to study the transition between intrawire and interwire dominated interactions, and further what effect the topology has on the transition. This is the subject of part III of this thesis. 

